First paper - Draft

### 2017-06-20 paper/paper.Rnw
First try - no changes
Plotted cor between dispersion and achievement - very high cor

@@

One important drawback of these studies is that they don't present the uncertainty of this estimate. Not necessarily to gauge their statistical significance, but to simply asses how much we can trust these numbers.

@@

Say more above Corak and Waldfogel above (book).

Despite their high level of rigurosity, their analysis is based on four surveys that have significant differences and cannot be easily compared. Their findings are very reliable but should be taken as suggestive.  For this reason we should also pay particular attention to studies such as Chmielewski and Reardon (2016) and Chmielewski (2017) which have attempted not only to compare gaps between countries, but to evaluate whether there is a general increase in educational inequality in many countries. These studies tackle a completely different question from the above, but they do provide support for the overall finding that the achievement gap is certainly not narrowing over time.

>>>>

<<>>=

country_scores <-
  map(adapted_year_data, ~ {
  if (unique(.x$wave) == "pisa2015") {
    intsvy::pisa2015.mean.pv("MATH", by = "country", data = .x)
  } else {
    intsvy::pisa.mean.pv("MATH", by = "country", data = .x)
  }
}) %>% setNames(seq(2000, 2015), 3) %>%

country_scores <-
  enframe(country_scores, name = "wave") %>%
  unnest() %>%
  mutate(wave = as.double(wave))

data_to_plot <-
  left_join(complete_data_topbottom, country_scores, by = c("wave", "country")) %>%
  select(wave, country, type_test, difference, Mean) %>%
  filter(type_test == "math")

data_to_plot %>%
  ggplot(aes(scale(Mean), difference)) +
  geom_point() +
  geom_smooth()

data_to_plot %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), difference, group = country)) +
  geom_point(colour = "blue") +
  geom_line(colour = "blue") +
  geom_point(aes(y = scale(Mean)), colour = "red") +
  geom_line(aes(y = scale(Mean)), colour = "red") +
  facet_wrap(~ country)

data_to_plot %>%
  semi_join(data_to_plot %>% group_by(country) %>% count() %>% filter(n > 3), by = "country") %>%
  group_by(country) %>%
  summarize(cor = cor(scale(Mean), difference, use = 'complete.obs')) %>%
  ggplot(aes(reorder(country, cor), cor, fill = cor > 0)) + geom_col() + coord_flip()

@

@@

One important drawback of these studies is that they don't present the uncertainty of this estimate. Not necessarily to gauge their statistical significance, but to simply asses how much we can trust their accuracy. Despite the recent surge in understanding achievement gaps as a function of social background, sociologists have already amassed a great deal of findings concerning inter-generational inequalities, one notable example being that it's very clear that large inequalities in parent's education tend to reproduce large inequalities in children's education (Breen & Jonsson 2007). Virtually in all countries, be it developed or developing, there is inequality of opportunity. But there is considerable variation in the magnitud of this effect. For example, the Scandinavian countries, particularly Denmark and Sweden, prove to be very mobile countries (Gosta and his phd student, Breen and other people, Shavit first study).

@@

One limitation of their study is that they concentrate on countries which have a very particular educational structure, namely the fact that there is little formal stratification in terms of curricula. These four countries have no major jump in school selection, quite the opposite to the European average countries. A thorough review by Van der Werfhorst and Mijs (2010) sheds some light on the subject. First and foremost, they gather substantive evidence showing that countries which have a highly tracked curriculum tend to have high levels of inequality, measured in terms of achievement gaps. Hanushek and Woessman (tracking study), use the TIMSS and PISA surveys to gauge whether highly tracked countries do indeed increase inequality after students pass the age at first selection of tracking. Moreover, Mickleweight & Schnepf (2007) using similar data but different design finds that countries which have a high level of tracking, are distinctively unequal in the difference between the top 95th and bottom 5th performers. In fact, the difference in test scores between these two groups is about 10 times higher then the average annual gain of a year of schooling.

Despite their high level of rigurosity, their analysis is based on four surveys that have significant differences in terms of questions, sampling and populations and cannot be easily compared. Their findings are very reliable but should be taken as suggestive at best.  For this reason we should also pay particular attention to studies such as Chmielewski and Reardon (2016) and Chmielewski (2017) which have attempted not only to compare gaps between countries, but to evaluate whether there is a general increase in educational inequality in many countries using similar studies. These studies tackle a completely different question from the above, but they do provide support for the overall finding that the achievement gap is certainly not narrowing over time.
